Everthing you wanted to know about PanDA but were truly truly afraid to ask

### 2016-12-05 overall.tex
finishing the write up for the I/O impact section

\subsection{PanDA Metadata Performance Impact at OLCF}

@@

When ATLAS was originally deployed at the OLCF, all shared libraries were
placed at the Spider 2 Lustre file system. However, it was quickly discovered
that the I/O pattern at the Athena stage was detrimental to the overall Spider
2 performance. Since Spider 2 is a center-wide file system and shared by all
OLCF resources and users, this resulted in lower overall interactive and
metadata performance at OLCF. As can be seen in Listing~\ref{mdstrace} the
metadata I/O activity for ATLAS exhibits a spike corresponding to the beginning
of the runs (i.e., the Athena stage) before tapering off.

\begin{lstlisting}[language=bash,frame=single,basicstyle=\ttfamily\tiny,caption=ATLAS metadata trace,label=mdstrace] 
XK7 Application 9205593
      39012 RPCs from 300 of 300 nodes
        ~69291.96 per sec
          37851 LDLM_ENQUEUE RPCs    ~67229.83 per sec
                pmin 13us pavg 42us pmax 4983us
          611 LDLM_CANCEL RPCs    ~1085.24 per sec
                pmin 10us pavg 16us pmax 32us
          277 MDS_CLOSE RPCs    ~492.00 per sec
                pmin 15us pavg 20us pmax 38us
          154 MDS_READPAGE RPCs    ~273.53 per sec
                pmin 170us pavg 292us pmax 671us
          86 MDS_GETXATTR RPCs    ~152.75 per sec
                pmin 15us pavg 21us pmax 65us
          30 MDS_GETATTR RPCs    ~53.29 per sec
                pmin 16us pavg 20us pmax 27us
          3 MDS_REINT RPCs    ~5.33 per sec
                pmin 103us pavg 136us pmax 196us
          Overall times
                pmin 10us pavg 42us pmax 4983us

@@

\end{lstlisting}

@@

\subsection{PanDA File I/O Performance Impact at OLCF}

File I/O performance and impact

@@

@@

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_file_close_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file close() histogram on Titan for week of 10/25/16.}
    \label{fig:atlas-titan_file_close}
\end{figure}

@@

>>>>

\subsection{PanDA Shared Library I/O Performance Impact at OLCF}

@@

When ATLAS was originally deployed at the OLCF, all shared libraries were
placed at the Spider 2 Lustre file system. However, it was quickly discovered
that the I/O pattern at the Athena stage was detrimental to the overall Spider
2 file system performance. Since Spider 2 is a center-wide file system and
shared by all OLCF resources and users, this resulted in lower overall
interactive and metadata performance at OLCF. As can be seen in
Listing~\ref{mdstrace} the metadata I/O activity for ATLAS exhibits a spike
corresponding to the beginning of the runs (i.e., the Athena stage) before
tapering off.

\begin{minipage}{\linewidth}
\begin{lstlisting}[language=bash,frame=single,basicstyle=\ttfamily\tiny,caption=ATLAS metadata trace,label=mdstrace] 
XK7 Application 9205593
      39012 RPCs from 300 of 300 nodes
        ~69291.96 per sec
          37851 LDLM_ENQUEUE RPCs    ~67229.83 per sec
                pmin 13us pavg 42us pmax 4983us
          611 LDLM_CANCEL RPCs    ~1085.24 per sec
                pmin 10us pavg 16us pmax 32us
          277 MDS_CLOSE RPCs    ~492.00 per sec
                pmin 15us pavg 20us pmax 38us
          154 MDS_READPAGE RPCs    ~273.53 per sec
                pmin 170us pavg 292us pmax 671us
          86 MDS_GETXATTR RPCs    ~152.75 per sec
                pmin 15us pavg 21us pmax 65us
          30 MDS_GETATTR RPCs    ~53.29 per sec
                pmin 16us pavg 20us pmax 27us
          3 MDS_REINT RPCs    ~5.33 per sec
                pmin 103us pavg 136us pmax 196us
          Overall times
                pmin 10us pavg 42us pmax 4983us

@@

\end{lstlisting}
\end{minipage}

@@

\subsection{PanDA I/O Impact at OLCF}

To better understand the I/O impact of ATLAS PanDA project on Titan
supercomputing environment we analyzed 1,175 jobs ran on the week of
10/25/2016, for a total of 174 hours. Table~\ref{panda-olcf-stats} shows the
overall statistical breakdown of the observed file I/O impact of ATLAS at OLCF.
Figures \ref{fig:atlas-titan-io-read} and \ref{fig:atlas-titan-io-written} show
the file read and write I/O histograms for these 1,175 jobs, respectively.
Figures \ref{fig:atlas-titan-file-open} and \ref{fig:atlas-titan-file-close}
show the file open() and close() metadata load histograms of the same 1,175
ATLAS jobs, respectively.

As can be seen from Table~\ref{panda-olcf-stats}, the number of nodes used by
ATLAS jobs vary between 1 and 300, while the average is at 35. 75\% of the
ATLAS jobs consume less than 25 and 92\% consume less than or equal to 100
Titan compute nodes. During the 174 hours of data collection, we observed that
6.75 ATLAS jobs were executed on average per hour on Titan and ran for
1.74 hours on average.

ATLAS jobs are file read I/O heavy, as can be seen in
Table~\ref{panda-olcf-stats}. The maximum amount read by any ATLAS job in
aggregate in this observed period was less than 250 GB and the maximum amount
written in aggregate was less than 75 GB. The average amount read per job is 20
GB and average amount written is 6 GB.

@@

Per job read and write file I/O statistics show an interesting pattern.
Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
compute node on Titan is less than 400 MB on average, while the amount of data
written per node is less than 170 MB on average. This correlates with our
finding that ATLAS PanDA jobs are read heavy. However, as can be seen in
Table~\ref{panda-olcf-stats} and figures {fig:atlas-titan-io-read} and
{fig:atlas-titan-io-written}, the distribution between read and written amount
of data per job is quite different from one another. The read I/O distribution
per job shows a long tail, ranging from 25 GB to 250 GB, while the written
amount of data has a very narrow distribution.

@@

\begin{figure}[!htb]
    \centering
    \begin{tabular}{cc}
        {\includegraphics[width=0.48\textwidth]{figures/panda_file_close_hist.pdf}}\\
    \end{tabular}
    \caption{ATLAS file close() histogram on Titan for week of 10/25/16.}
    \label{fig:atlas-titan-file-close}
\end{figure}

On the metadata I/O breakdown, ATLAS PanDA jobs yield 23 file open() operations
and 5 file close() operations per second. As can be seen from figures
\ref{fig:atlas-titan-file-open} and \ref{fig:atlas-titan-file-close}, they
exhibit similar distributions. The maximum number of file open() operations are
around 170 on average and the maximum number of file close() operations are 39
on average per second per job. The total number of file open() operations is
172,089,760 for the observed window of 174 hours, while the total number of
file close() operations stand at 40,132,992 for the total 1,175 ATLAS PanDA
jobs in the same observation window. The difference between these two values is
puzzling and it is under investigation at the time of writing this paper.
{\sarp{can someone take a look at this? I would expect at least similar values
for number of opens and closes and there is vast difference. Can it be that
ATLAS jobs do a lot of opens without following closes, or perhaps multiple
opens for the same files over and over?}}

@@

Overall, based on our experiments with real-world jobs, it can be safely
concluded that the file and metadata I/O load of ATLAS PanDA project on the
OLCF Titan supercomputing environment and the Spider 2 file system is not
detrimental to the center operations and the overall impact is minimal, at the
current scale of the project.
